---
title: "Experimental model-derived EdU report (non-production)"
format: { html: { toc: true, embed-resources: true } }
execute: { warning: false, message: false }
params: { config: null, analysis_rds: null, model_gate_manifest: null }
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
model_profile <- analysis$config$gating$profile
model_manifest_path <- params$model_gate_manifest
if (is.null(model_manifest_path) && identical(model_profile, "documented_edu_g1_v1")) {
  model_manifest_path <- analysis$config$gating$manifest
}
if (is.null(model_manifest_path)) stop("`model_gate_manifest` is required.")
model_gate_inputs <- validate_edu_model_gate_manifest(analysis, model_manifest_path)
report <- build_edu_pseudocolor_output_contract(analysis)
```

> **EXPERIMENTAL MODEL-DERIVED NON-PRODUCTION.** Single Cells, G1, and EdU+
> use the explicitly configured model profile `r model_profile`. No FlowJo gate
> is substituted. This report
> does not establish biological ground truth, production performance, or broad
> generalizability.

The report validates the selected profile's immutable model identities,
thresholds, feature order, event identities, and provenance. It does not mix
artifacts or thresholds from the other available model profile.

## Every-sample DNA-versus-EdU panels

```{r}
for (prefix in names(report$panels)) print(report$panels[[prefix]])
```

## EdU-positive percentages

EdU positivity uses the selected profile's approved classifier and is
intersected by direct event identity with each model-derived Single Cells
population.

```{r}
print(report$positivity$panels$two_to_four_n)
print(report$positivity$panels$early_mid_late_s)
```

## EdU-positive intensity

```{r}
print(report$intensity$panels$all_computed_positive)
print(report$intensity$panels$early_mid_late_s)
```

## Model-gating provenance

```{r}
print(model_gate_inputs)
print(attr(model_gate_inputs, "model_artifacts"))
```

Both profiles are experimental and retain their separately approved G1
definitions. `single_g1_edu_frozen_v1` excludes model-called EdU+ events before
background calculation and applies its acquisition-relative DNA minimum;
`documented_edu_g1_v1` retains the documented G1 model and 0.29 threshold.
These definitions are not combined. Both operate outside their original
expert-gated parent distributions and must not be represented as biological
ground truth or production gating.

## Software used

```{r}
facs_report_software_used()
```
