---
title: "EdU every-sample pseudocolor report"
format: { html: { toc: true, embed-resources: true } }
execute: { warning: false, message: false }
params:
  config: null
  analysis_rds: null
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
report <- build_edu_pseudocolor_output_contract(analysis)
```

## Every-sample DNA-versus-EdU panels

```{r}
for (prefix in names(report$panels)) print(report$panels[[prefix]])
```

## EdU-positive percentages

The 2N–4N panel uses all positivity-eligible events with DNA content from 2N
through 4N, inclusive, as its denominator. The regional panel uses the
established canonical Early/Mid/Late-S denominators; it is therefore reported
separately rather than labeled “All S”.

```{r}
print(report$positivity$panels$two_to_four_n)
print(report$positivity$panels$early_mid_late_s)
```

The positivity source tables are retained in the completed analysis artifact
and are intentionally not printed in this reader-facing report.

## EdU-positive intensity

These figures use only the completed canonical background-subtracted EdU
intensity tables. Within each biological replicate, each value is divided by
the explicitly configured matched Untreated value; Untreated is therefore 1.
They do not re-fit background, re-call positivity, or change any event-level
value. The percentage figures above remain absolute percentages.

When a completed background-subtracted Untreated median is nonpositive or a
canonical median is nonfinite, its reference-relative value is shown as
unavailable rather than replaced by a direct, raw, or display-offset value.

```{r}
print(report$intensity$panels$all_computed_positive)
print(report$intensity$panels$early_mid_late_s)
```

The intensity source tables are retained in the completed analysis artifact
and are intentionally not printed in this reader-facing report.

## Display provenance

Display-offset, panel-QC, and provenance records remain in the completed
analysis artifact for audit and are not printed here as source tables.

## Software used

This report was rendered with Quarto. The versions below identify the analysis
and report-rendering software used for this report.

```{r}
facs_report_software_used()
```
