---
title: "EdU experiment report"
subtitle: "FlowJo populations, quality checks, cell-cycle views, and quantitative results"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: false
    embed-resources: true
    code-fold: false
    code-tools: false
execute: { warning: false, message: false, echo: false }
params:
  config: null
  analysis_rds: null
  show_audit_tables: false
  max_condition_columns: 4
  min_panel_width: 220
---

```{css}
.quarto-title-block { margin-bottom: 2rem; }
h2 { margin-top: 2.4rem; padding-top: .45rem; border-top: 2px solid #dce5e8; }
h3 { margin-top: 1.7rem; }
.report-card { border-left: .35rem solid #287c8e; background: #f5f9fa; padding: 1rem 1.2rem; margin: 1rem 0; }
.report-alert { border-left-color: #b54708; background: #fff7ed; }
.report-ok { border-left-color: #27864b; background: #f0f9f3; }
.report-table { width: 100%; font-size: .9rem; }
.report-table th { background: #eaf1f3; }
.figure-caption, figcaption { color: #46565c; font-size: .9rem; }
.overview-grid { display: grid; grid-template-columns: repeat(auto-fit, minmax(min(100%, var(--panel-min)), 1fr)); gap: 1rem; max-width: var(--panel-max); }
.overview-card { min-width: 0; border: 1px solid #dce5e8; border-radius: .35rem; padding: .55rem; background: #fff; }
.overview-card img { width: 100%; height: auto; }
.overview-card h4 { margin: .5rem 0 .15rem; font-size: .95rem; }
.overview-card .sample-caption { margin: 0; color: #46565c; font-size: .82rem; overflow-wrap: anywhere; }
.overview-group { margin: 1.5rem 0 2.2rem; }
@media (prefers-reduced-motion: reduce) { * { scroll-behavior: auto !important; } }
```

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
report <- build_edu_pseudocolor_output_contract(analysis)
overview <- facs_report_edu_overview(analysis)
qc_sections <- facs_report_edu_qc_sections(analysis, report)
all_conditions <- facs_report_edu_responsive_groups(
  analysis, report, params$max_condition_columns, params$min_panel_width
)
```

```{r}
#| label: pre-results-qc
#| results: asis
pre_results_qc <- rbind(
  qc_sections$data_input,
  qc_sections$normalization
)
pre_results_findings <- pre_results_qc[pre_results_qc$level != "CLEAR", , drop = FALSE]
if (nrow(pre_results_findings)) {
  cat('<div class="report-card report-alert"><strong>Pre-results QC: review required.</strong> One or more retained data, input, normalization, or quantitation states are non-clear. Review the findings below before interpreting any plot.</div>\n')
  cat(facs_report_compact_table(pre_results_findings))
} else {
  cat('<div class="report-card report-ok"><strong>Pre-results QC: retained checks are clear.</strong> Human inspection of every acquisition is still required before interpretation.</div>\n')
}
```

## All-condition visual overview

```{r}
#| label: all-condition-matrix
#| results: asis
for (group in all_conditions$groups) {
  cat('\n\n::: {.overview-group}\n')
  cat("<h3>", facs_report_heading_label(group$replicate),
      " · technical acquisition ",
      facs_report_heading_label(group$technical_replicate),
      "</h3>\n", sep = "")
  cat("<p><strong>Paired acquisition group:</strong> ",
      facs_report_heading_label(group$model_group),
      " · <strong>Configured reference:</strong> ",
      facs_report_heading_label(group$reference_condition), "</p>\n", sep = "")
  if (!identical(group$legend_status, "available")) {
    cat('<div class="report-card report-alert"><strong>Density legend unavailable:</strong> no available panel in this paired acquisition group supplied a legend. Every card remains represented.</div>\n')
  }
  maximum_width <- all_conditions$max_condition_columns *
    all_conditions$min_panel_width + 16 * (all_conditions$max_condition_columns - 1L)
  cat(sprintf(
    '::: {.overview-grid style="--panel-min: %.0fpx; --panel-max: %.0fpx;"}\n',
    all_conditions$min_panel_width, maximum_width
  ))
  for (card_index in seq_along(group$cards)) {
    card <- group$cards[[card_index]]
    card_title <- if (isTRUE(card$is_reference)) {
      paste0(card$condition, " — configured reference")
    } else card$condition
    caption_id <- paste0("overview-card-", group$group_index, "-", card_index)
    cat('::: {.overview-card role="group" aria-labelledby="', caption_id, '"}\n', sep = "")
    cat(facs_report_knit_plot(
      card$plot, width = 2.3, height = 2.15,
      envir = environment(), dpi = 192L
    ), sep = "\n")
    cat('<h4 id="', caption_id, '">',
        facs_report_heading_label(card_title), '</h4>\n', sep = "")
    if (!is.na(card$prefix)) {
      cat('<p class="sample-caption">Sample: ',
          facs_report_heading_label(card$prefix), '</p>\n', sep = "")
    }
    cat(':::\n')
  }
  cat(':::\n')
  if (!is.null(group$legend)) {
    cat(facs_report_knit_plot(group$legend, width = 7, height = .7,
                              envir = environment()), sep = "\n")
  }
  cat(':::\n')
}
```

## Quantitation overview

```{r}
#| label: quantitation-overview
#| results: asis
quant_point_alt <- if (isTRUE(analysis$config$quant_show_points)) {
  "Points are biological-replicate values. "
} else {
  "Biological-replicate points are disabled by configuration. "
}
quant_error_alt <- if (identical(analysis$config$quant_error_bar, "none")) {
  "Error bars are disabled by configuration."
} else {
  paste0("Error bars use the configured ", analysis$config$quant_error_bar,
         " display.")
}
quantitation_rows <- list(
  list(
    list(
      title = "2N–4N EdU-positive percentage",
      detail = "Absolute EdU-positive percentage among eligible Single Cell events in the inclusive 2N–4N interval",
      alt = paste0("EdU-positive percentage among eligible Single Cell events in the inclusive 2N-to-4N DNA interval. ", quant_point_alt, "Bar heights are condition means. ", quant_error_alt),
      y_label = "% EdU positive",
      small_legend_keys = TRUE,
      hide_single_category_x_tick = TRUE,
      hide_x_title = TRUE,
      plot = report$positivity$panels$two_to_four_n
    ),
    list(
      title = "All computed-positive EdU intensity",
      detail = "Established unweighted mean of technical-acquisition background-subtracted medians among all computed-positive eligible cells within each biological replicate, relative to its configured matched reference",
      alt = paste0("EdU intensity for all computed-positive eligible cells, relative to each biological replicate's configured matched reference. ", quant_point_alt, "Horizontal marks are condition means. ", quant_error_alt),
      y_label = "Relative EdU level",
      small_legend_keys = FALSE,
      hide_single_category_x_tick = TRUE,
      hide_x_title = TRUE,
      plot = report$intensity$panels$all_computed_positive
    )
  ),
  list(
    list(
      title = "Regional S-phase percentage",
      detail = "Absolute percentage within each established Early/Mid/Late-S DNA region",
      alt = paste0("EdU-positive percentage within the established Early, Mid, and Late S-phase DNA regions. ", quant_point_alt, "Bar heights are condition means. ", quant_error_alt),
      y_label = "% EdU positive",
      small_legend_keys = TRUE,
      hide_single_category_x_tick = FALSE,
      hide_x_title = FALSE,
      plot = report$positivity$panels$early_mid_late_s
    ),
    list(
      title = "Regional S-phase EdU",
      detail = "Established unweighted mean of technical-acquisition background-subtracted medians within each biological replicate and established S-phase region, relative to its configured matched reference",
      alt = paste0("EdU intensity within the established Early, Mid, and Late S-phase regions, relative to each biological replicate's configured matched reference. ", quant_point_alt, "Horizontal marks are condition means. ", quant_error_alt),
      y_label = "Relative EdU level",
      small_legend_keys = FALSE,
      hide_single_category_x_tick = FALSE,
      hide_x_title = FALSE,
      plot = report$intensity$panels$early_mid_late_s
    )
  )
)
quantitation_cards <- unlist(quantitation_rows, recursive = FALSE)
quantitation_row_categories <- c(1L, 3L)
quantitation_card_ids <- facs_report_card_ids(
  "quantitation-card", length(quantitation_cards)
)
quantitation_condition_count <- length(unique(as.character(
  analysis$sample_manifest$condition
)))
quantitation_legend <- facs_report_quantitation_legend_layout(
  quantitation_condition_count
)
card_index <- 0L
for (row_index in seq_along(quantitation_rows)) {
  row <- quantitation_rows[[row_index]]
  card_width <- facs_report_quantitation_card_width(
    quantitation_condition_count,
    quantitation_row_categories[[row_index]]
  )
  plot_width <- facs_report_quantitation_plot_width(card_width)
  row_layout <- facs_report_quantitation_row_layout(
    length(row), card_width
  )
  cat(sprintf(
    '::: {.overview-grid .quantitation-row style="--panel-min: %.0fpx; --panel-max: %.0fpx;"}\n',
    row_layout$min_panel_width, row_layout$maximum_width
  ))
  for (card in row) {
    card_index <- card_index + 1L
    caption_id <- quantitation_card_ids[[card_index]]
    cat('::: {.overview-card role="group" aria-labelledby="', caption_id, '"}\n', sep = "")
    compact_plot <- card$plot +
      ggplot2::labs(
        title = NULL, subtitle = NULL, caption = NULL, y = card$y_label,
        fill = NULL, colour = NULL
      ) +
      ggplot2::theme(
        legend.position = "bottom", legend.box = "vertical",
        legend.box.just = "left"
      ) +
      facs_report_compact_typography(
        title_size = 8, secondary_size = 7.5, legend_size = 7
      ) +
      ggplot2::theme(legend.title = ggplot2::element_blank()) +
      ggplot2::guides(
        fill = ggplot2::guide_legend(
          title = NULL, ncol = quantitation_legend$columns, byrow = TRUE
        ),
        colour = ggplot2::guide_legend(
          title = NULL, ncol = quantitation_legend$columns, byrow = TRUE
        )
      )
    if (isTRUE(card$small_legend_keys)) {
      compact_plot <- compact_plot +
        facs_report_compact_legend_key_theme(0.28)
    }
    if (isTRUE(card$hide_single_category_x_tick)) {
      compact_plot <- compact_plot + ggplot2::theme(
        axis.text.x = ggplot2::element_blank(),
        axis.ticks.x = ggplot2::element_blank()
      )
    }
    if (isTRUE(card$hide_x_title)) {
      compact_plot <- compact_plot + ggplot2::theme(
        axis.title.x = ggplot2::element_blank()
      )
    }
    cat(facs_report_knit_plot(
      compact_plot,
      width = plot_width, height = quantitation_legend$figure_height,
      envir = environment(), dpi = 192L,
      fig_alt = card$alt
    ), sep = "\n")
    cat('<h4 id="', caption_id, '">',
        facs_report_heading_label(card$title), '</h4>\n', sep = "")
    cat('<p class="sample-caption">',
        facs_report_heading_label(card$detail), '</p>\n', sep = "")
    cat(':::\n')
  }
  cat(':::\n')
}
```

## Experiment overview

::: {.report-card}
This report follows the configured sample order and uses the established
FlowJo-derived **Single Cells**, **G1**, and **EdU Positive** populations. It
does not redraw gates or recalculate phase boundaries. The opening galleries
provide rapid orientation; use this experiment summary and the pre-results QC
status before interpreting the detailed results.
:::

```{r}
#| label: overview-summary
#| results: asis
cat(facs_report_compact_table(overview$summary))
```

### Samples and replicate structure

```{r}
#| label: overview-samples
#| results: asis
cat(facs_report_compact_table(overview$samples))
```

## Every-sample DNA-versus-EdU plots

Each panel represents one named acquisition. Color encodes local event density,
not EdU amount. The y-axis uses the established background-subtracted signal
plus a display-only, per-sample offset so it can be shown on the configured
scale; the offset is never used for positivity or quantitative medians.

```{r}
#| label: every-sample-panels
#| results: asis
#| fig-cap: "DNA content versus EdU for every configured acquisition. Panel titles and subtitles preserve the sample, condition, biological replicate, and technical-acquisition identity."
for (prefix in names(report$panels)) {
  cat("\n\n<h3>", facs_report_heading_label(prefix), "</h3>\n\n", sep = "")
  print(report$panels[[prefix]])
}
```

## S-phase views

Early, Mid, and Late S use the established DNA regions and their established
regional denominators. These categories are shown separately to avoid implying
that their denominator is the same as the inclusive 2N–4N result.

```{r}
#| label: s-phase-positivity
#| fig-cap: "EdU-positive percentage within each established S-phase DNA region. Points are biological replicates; bars are condition means."
print(report$positivity$panels$early_mid_late_s)
```

## EdU-positive percentages

The 2N–4N result uses all positivity-eligible events with DNA content from 2N
through 4N, inclusive, as its denominator. It is an absolute percentage and is
not reference-normalized.

```{r}
#| label: two-to-four-n-positivity
#| fig-cap: "Computed EdU-positive events as a percentage of positivity-eligible Single Cell events in the inclusive 2N–4N DNA interval."
print(report$positivity$panels$two_to_four_n)
```

## Background-subtracted EdU measurements

These figures use only the completed canonical background-subtracted EdU
intensity tables. Within each biological replicate, each value is divided by
the explicitly configured matched reference value; that reference is therefore 1.
They do not re-fit background, re-call positivity, or change any event-level
value. The percentage figures above remain absolute percentages.

When a completed background-subtracted matched-reference median is nonpositive
or a canonical median is nonfinite, its reference-relative value is shown as
unavailable rather than replaced by a direct, raw, or display-offset value.

```{r}
#| label: positive-intensity-overall
#| fig-cap: "Median background-subtracted EdU among all computed EdU-positive eligible cells, relative to the matched configured reference within each biological replicate."
print(report$intensity$panels$all_computed_positive)
```

```{r}
#| label: positive-intensity-regional
#| fig-cap: "Median background-subtracted EdU among computed EdU-positive eligible cells in each established S-phase region, relative to the matched configured reference."
print(report$intensity$panels$early_mid_late_s)
```

## Methods

- **Population inputs:** configured FlowJo Single Cells, G1, and EdU Positive
  exports. The report does not redefine these gates.
- **DNA coordinate:** each acquisition's G1-derived anchor, calculated using
  the configured anchor method, is mapped to 2N; displayed 4N is twice that
  value.
- **EdU correction:** an independent EdU-negative baseline slope is fitted for
  each acquisition and combined with that acquisition's G1-derived anchor using
  the configured anchor method to form that acquisition's baseline, which is
  subtracted. The event plots add a display-only offset; quantitative medians
  use unoffset background-subtracted values.
- **Percentages:** each percentage retains its named canonical numerator and
  denominator. The 2N–4N interval is inclusive; regional S-phase denominators
  remain separate.
- **Replicates:** technical acquisitions are combined by the established
  unweighted mean within biological replicate. Plot points represent biological
  replicates; condition marks or bars summarize them using the configured error
  display.
- **Reference-relative intensity:** canonical biological-replicate medians are
  divided by the explicitly configured matched reference. Invalid denominators
  remain unavailable rather than being replaced.

## Provenance

### Share-ready provenance

This summary uses file names and configured identities only; it intentionally
omits directory paths.

```{r}
#| label: share-provenance
#| results: asis
share_inputs <- unique(data.frame(
  `Input file` = basename(as.character(overview$inputs$path)),
  Population = as.character(overview$inputs$population),
  check.names = FALSE, stringsAsFactors = FALSE
))
share_summary <- data.frame(
  Item = c("Configuration", "Acquisitions", "Biological replicates", "Conditions", "Input files represented"),
  Value = c(
    overview$summary$value[overview$summary$item == "Configuration"],
    nrow(analysis$sample_manifest), length(unique(analysis$sample_manifest$replicate_index)),
    length(unique(analysis$sample_manifest$condition)), nrow(share_inputs)
  ), stringsAsFactors = FALSE
)
cat(facs_report_compact_table(share_summary))
cat("\n\n<details><summary>Named input files (paths omitted)</summary>\n\n")
cat(facs_report_compact_table(share_inputs))
cat("\n\n</details>")
```

```{r}
#| label: provenance
#| results: asis
if (isTRUE(params$show_audit_tables)) {
  cat("**Path disclosure enabled:** full validated input paths are shown because `show_audit_tables: true` was explicitly requested. Review the report before sharing it.\n\n")
  cat("### Display-offset audit\n\n")
  cat(facs_report_compact_table(report$display_offset_qc))
  cat("\n\n### Panel audit\n\n")
  cat(facs_report_compact_table(report$panel_qc))
  cat("\n\n### Named input files\n\n")
  cat(facs_report_compact_table(overview$inputs))
} else {
  cat("Detailed audit tables are hidden by report configuration; they remain in the completed analysis artifact.")
}
```

### Software used

This report was rendered with Quarto. The versions below identify the analysis
and report-rendering software used for this report.

```{r}
facs_report_software_used()
```
