---
title: "pH3 condition report"
format: { html: { toc: true, embed-resources: true } }
execute: { message: false }
params:
  config: null
  analysis_rds: null
  appearance_file: null
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"),
       local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
if (!identical(analysis$config$plot_type, "ph3") ||
    !inherits(analysis$ph3_output_model, "ph3_output_model")) {
  stop("This report requires a completed production pH3 output-contract analysis.",
       call. = FALSE)
}
plots <- plot_ph3_condition_report(
  analysis,
  appearance = facs_report_appearance(params),
  appearance_file = params$appearance_file
)
report_model <- analysis$ph3_output_model$condition_report
```

## Required pH3 outcomes

Each panel shows the retained biological-replicate values and the condition
mean. A standard deviation is shown only when at least three finite biological
replicate values are available. No inferential statistics are shown in this
profile.

### 4N pH3-positive prevalence

```{r}
plots$panels$ph3_4n_positive_prevalence
```

### Below-4N pH3-positive prevalence

```{r}
plots$panels$ph3_below_4n_positive_prevalence
```

### pH3 signal in 4N pH3-positive cells

```{r}
plots$panels$ph3_4n_positive_signal
```

### pH3 signal in below-4N pH3-positive cells

```{r}
plots$panels$ph3_below_4n_positive_signal
```

## Concise QC and provenance

The tables below preserve availability, partial-coverage, and mixed-basis
states. A mixed signal basis is never combined into one condition signal mean.

```{r}
report_model$qc_flags
```

```{r}
report_model$provenance
```

```{r}
analysis$ph3_output_model$correction
analysis$ph3_output_model$geometry
```

## Computed positivity-cutoff QC

When the owner-confirmed raw-4N density method is configured, each explicit
biological-replicate set uses its matched untreated control's raw-pH3 4N
density. The dashed line is the frozen cutoff applied unchanged to that
control and its matched treatment(s). FlowJo pH3-positive membership is
retained only as source provenance and is not the numerical cutoff authority.

```{r}
if (length(plots$cutoff_qc)) {
  for (panel in plots$cutoff_qc) print(panel)
  plots$cutoff_records
  plots$cutoff_application_qc
} else {
  cat("Legacy FlowJo positivity mode: no computed raw-4N cutoff was configured.\n")
}
```

```{r}
#| label: analysis-warnings
#| results: asis
if (length(analysis$warnings)) {
  cat("### Analysis warnings\n\n")
  cat(paste0("- ", analysis$warnings), sep = "\n")
  cat("\n")
}
```
