---
title: "Experimental model-derived EdU report (non-production)"
format: { html: { toc: true, embed-resources: true } }
execute: { warning: false, message: false }
params: { config: null, analysis_rds: null, model_gate_manifest: null }
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
if (is.null(params$model_gate_manifest)) stop("`model_gate_manifest` is required.")
analysis <- facs_report_input(params$config, params$analysis_rds)
model_gate_inputs <- validate_edu_model_gate_manifest(analysis, params$model_gate_manifest)
report <- build_edu_pseudocolor_output_contract(analysis)
```

> **EXPERIMENTAL MODEL-DERIVED NON-PRODUCTION.** Single Cells and G1 were
> called with immutable development-only models, not FlowJo gates. This report
> does not establish biological ground truth, production performance, or broad
> generalizability.

## Every-sample DNA-versus-EdU panels

```{r}
for (prefix in names(report$panels)) print(report$panels[[prefix]])
```

## EdU-positive percentages

EdU positivity remains the explicitly mapped FlowJo annotation intersected by
direct event identity with each model-derived Single Cells population.

```{r}
print(report$positivity$panels$two_to_four_n)
print(report$positivity$panels$early_mid_late_s)
```

## EdU-positive intensity

```{r}
print(report$intensity$panels$all_computed_positive)
print(report$intensity$panels$early_mid_late_s)
```

## Model-gating provenance

```{r}
print(model_gate_inputs)
print(attr(model_gate_inputs, "model_artifacts"))
```

The G1 model was developed within expert-reference Single Cells. Applying it
after model-derived Single Cells changes its parent distribution; prior G1
validation metrics therefore do not validate this composed use. The older G1
rule also omits the second-marker axes used by the audited expert G1 gates.

## Software used

```{r}
facs_report_software_used()
```
