---
title: "CCNA2-positive signal report — v1"
subtitle: "Background-corrected CCNA2 distributions in retained Single Cells"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    code-fold: false
execute: { warning: false, message: false, echo: false }
params:
  config: null
  analysis_rds: null
  positive_cutoff: 100000
  display_quantiles: [0.01, 0.99]
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
report <- facs_report_poi_positive_signal(
  analysis,
  positive_cutoff = params$positive_cutoff,
  display_quantiles = unlist(params$display_quantiles, use.names = FALSE)
)
target_name <- analysis$config$target_name
```

## Definition and display scope

```{r}
#| results: asis
cat('<div class="callout-note"><p><strong>CCNA2-positive definition:</strong> retained Single Cell events with background-corrected ',
    target_name, ' at or above <code>', format(report$positive_cutoff, scientific = TRUE),
    '</code>. This is a report-specific threshold; it does not alter the saved FlowJo populations, the background model, or the existing POI report.</p></div>\n', sep = '')
cat('<div class="callout-note"><p><strong>Display range:</strong> the y-axis viewport is the ',
    report$display_quantiles[[1L]] * 100, 'th through ', report$display_quantiles[[2L]] * 100,
    'th percentile of all finite retained Single Cell values (',
    format(report$display_limits[[1L]], big.mark = ','), ' to ',
    format(report$display_limits[[2L]], big.mark = ','),
    '). Values outside the viewport remain included in the violin and boxplot calculations.</p></div>\n', sep = '')
```

## Event-level CCNA2 distributions

```{r}
#| fig-width: 10
#| fig-height: 5.8
#| fig-alt: "Background-corrected CCNA2 violins and transparent boxplots for all retained Single Cells and CCNA2-positive cells, grouped by condition."
print(report$plot)
```

## Per-sample CCNA2-positive summary

```{r}
knitr::kable(
  report$sample_summary,
  digits = c(0, 0, 0, 0, 0, 2, 0),
  caption = "Finite retained Single Cells and report-defined CCNA2-positive events by sample."
)
```

## Methods

- **Signal:** background-corrected `r target_name` (`target_bgsub`), calculated by the existing POI analysis.
- **Positive threshold:** `r format(report$positive_cutoff, scientific = TRUE)` on the background-corrected signal.
- **Event distributions:** violins and transparent boxplots use all finite retained Single Cell event values; the positive group is the subset at or above the threshold.
- **Display bounds:** `r report$display_quantiles[[1L]] * 100`th–`r report$display_quantiles[[2L]] * 100`th percentiles are a coordinate viewport only and do not exclude observations from the plotted distribution statistics.
