---
title: "pH3 legacy CSV pilot report — NOT PUBLICATION-GRADE"
format: { html: { toc: true, embed-resources: true } }
execute: { message: false }
params:
  config: null
  analysis_rds: null
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
if (!identical(analysis$config$ph3_input_profile, "legacy_csv_pilot_v1") ||
    is.null(analysis$ph3_legacy_pilot)) {
  stop("This report requires a completed legacy_csv_pilot_v1 pH3 analysis.", call. = FALSE)
}
pilot <- plot_ph3_legacy_pilot_report(analysis)
```

## PILOT / LIMITED-PROVENANCE — NOT PUBLICATION-GRADE

This report reads existing historical FlowJo CSV exports. Event identity and
gate containment cannot be verified. Each sample receives an auditable
provisional raw-4N partition only to fit its own DNA-dependent linear
background model. Final cutoff, positivity, signal panels, and C/D outcomes
use **background-subtracted pH3**; the final cutoff is derived from the
matched Untreated sample on that corrected scale. For each explicit clone
group, pseudocolor panels use a display-only target equal to the median of that
group's per-sample raw final-negative medians. Each sample's offset is this
shared target minus its corrected final-negative median, so usable samples in
the group have aligned displayed negative medians. The offset does not alter
any cutoff, outcome, or summary. If any configured group member lacks a valid
final-negative display input, every panel in that group is shown as unavailable
rather than using a partial target. Panels then use a log10 y-axis. Any restored
display values at or below zero are excluded from the plotted density only and
are reported per sample; they remain unchanged in all stored analysis values.

The complete pilot limitations, source tables, cutoff records, QC records, and
provenance remain in the completed analysis artifact. They are intentionally
not printed in this reader-facing report.

## Four pilot outcomes

```{r}
pilot$panels$ph3_4n_positive_prevalence
pilot$panels$ph3_below_4n_positive_prevalence
pilot$panels$ph3_4n_positive_signal
pilot$panels$ph3_below_4n_positive_signal
```

## Background-subtracted pH3-versus-DNA pseudocolor QC

One display-only panel is shown for every explicitly configured pilot sample.
Panels with an available correction use background-subtracted pH3 from the
same complete-event tables, show the clone-matched final corrected 4N cutoff
when its restored display value is positive, and shade the below-4N and 4N analytic regions. A sample with
an unavailable correction instead shows a clearly labelled placeholder with
its stored correction status and reason; it shows no raw events, corrected
density, or fabricated cutoff. The visual DNA window is 1.6N--4.4N; the
analytical 2N--4N bounds remain visibly marked and are unchanged for available
panels. These are computed display regions, not reconstructed or verified
FlowJo geometry.

The display-offset, log10-display, cutoff, provisional-partition, and
background-regression QC records are retained with the completed analysis for
audit. They are not printed here as source tables.

```{r}
#| fig-width: 3.125
#| fig-height: 3.375
#| out-width: "3.125in"
#| fig-align: "left"
#| dpi: 144
for (prefix in names(pilot$pseudocolor_panels)) {
  print(pilot$pseudocolor_panels[[prefix]])
}
```

## Software used

This report was rendered with Quarto. The versions below identify the analysis
and report-rendering software used for this report.

```{r}
facs_report_software_used()
```
