---
title: "pH3 legacy CSV pilot report — NOT PUBLICATION-GRADE"
format: { html: { toc: true, embed-resources: true } }
execute: { message: false }
params:
  config: null
  analysis_rds: null
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
if (!identical(analysis$config$ph3_input_profile, "legacy_csv_pilot_v1") ||
    is.null(analysis$ph3_legacy_pilot)) {
  stop("This report requires a completed legacy_csv_pilot_v1 pH3 analysis.", call. = FALSE)
}
pilot <- plot_ph3_legacy_pilot_report(analysis)
```

## PILOT / LIMITED-PROVENANCE — NOT PUBLICATION-GRADE

This report reads existing historical FlowJo CSV exports. Event identity and
gate containment cannot be verified. Each sample receives an auditable
provisional raw-4N partition only to fit its own DNA-dependent linear
background model. Final cutoff, positivity, signal panels, and C/D outcomes
use **background-subtracted pH3**; the final cutoff is derived from the
matched Untreated sample on that corrected scale. For each explicit clone
group, pseudocolor panels use a display-only target equal to the median of that
group's per-sample raw final-negative medians. Each sample's offset is this
shared target minus its corrected final-negative median, so usable samples in
the group have aligned displayed negative medians. The offset does not alter
any cutoff, outcome, or summary. If any configured group member lacks a valid
final-negative display input, every panel in that group is shown as unavailable
rather than using a partial target. Panels then use a log10 y-axis. Any restored
display values at or below zero are excluded from the plotted density only and
are reported per sample; they remain unchanged in all stored analysis values.

```{r}
pilot$limitations
```

## Four pilot outcomes

```{r}
pilot$panels$ph3_4n_positive_prevalence
pilot$panels$ph3_below_4n_positive_prevalence
pilot$panels$ph3_4n_positive_signal
pilot$panels$ph3_below_4n_positive_signal
```

## Final corrected-cutoff QC

```{r}
pilot$cutoff_records
```

## Provisional raw partition and background-regression QC

These raw-scale cutoffs are audit intermediates only. They are not the final
positivity calls and do not appear as final panel overlays.

```{r}
pilot$provisional_background_qc
pilot$provisional_density_curves
```

## Background-subtracted pH3-versus-DNA pseudocolor QC

One display-only panel is shown for every explicitly configured pilot sample.
Panels with an available correction use background-subtracted pH3 from the
same complete-event tables, show the clone-matched final corrected 4N cutoff
when its restored display value is positive, and shade the below-4N and 4N analytic regions. A sample with
an unavailable correction instead shows a clearly labelled placeholder with
its stored correction status and reason; it shows no raw events, corrected
density, or fabricated cutoff. The visual DNA window is 1.6N--4.4N; the
analytical 2N--4N bounds remain visibly marked and are unchanged for available
panels. These are computed display regions, not reconstructed or verified
FlowJo geometry.

## Display-offset QC

For each available sample, the display-only offset is `median(raw final
negative) - median(background-subtracted final negative)`, calculated from the
same eligible final pH3-negative events used only for visual restoration. The
shown cutoff is translated by that same offset. This table is audit information;
all numerical outcomes above remain on the unshifted background-subtracted
scale.

```{r}
pilot$display_offset_qc
```

## Log10-display QC

The log10 axis is display-only. This table records finite events in each
visual DNA window, how many are positive and therefore displayed, and how many
nonpositive restored values are excluded from the plotted density. If no
positive-domain display events exist, the report shows an explicit no-data
panel and no cutoff overlay; it never falls back to raw pH3.

```{r}
pilot$log_display_qc
```

```{r}
#| fig-width: 3.125
#| fig-height: 3.375
#| out-width: "3.125in"
#| fig-align: "left"
#| dpi: 144
for (prefix in names(pilot$pseudocolor_panels)) {
  print(pilot$pseudocolor_panels[[prefix]])
}
```

```{r}
pilot$condition_summary
```
