---
title: "FlowJo-defined 4N pH3-positive populations"
format: { html: { toc: true, embed-resources: true } }
execute: { warning: false, message: false }
params:
  config: null
  analysis_rds: null
  gate_geometry_csv: null
  appearance_file: null
  output_analysis_rds: null
  output_figure_bundle_rds: null
  overwrite: false
  target_axis_label: null
  dna_axis_label: null
  palette: null
  x_limits: null
  y_limits: null
  y_log10: null
  point_size: null
  density_bandwidth: null
  density_lower_clip: null
  density_upper_clip: null
  density_gamma: null
  gate_color: null
  gate_linetype: null
  gate_linewidth: null
  gate_fill: null
  gate_alpha: null
  label_color: null
  label_size: null
  label_digits: null
  gate_top_inset_fraction: null
  layout_options: null
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"),
       local = TRUE)
analysis <- facs_report_input(params$config, params$analysis_rds)
if (!is.character(params$gate_geometry_csv) ||
    length(params$gate_geometry_csv) != 1L ||
    !nzchar(params$gate_geometry_csv)) {
  stop("Supply one explicit `gate_geometry_csv` path.", call. = FALSE)
}
geometry <- read_ph3_gate_geometry(params$gate_geometry_csv, analysis)
value_or <- function(value, default) if (is.null(value)) default else value
bundle <- build_ph3_4n_figure_bundle(
  analysis = analysis,
  geometry = geometry,
  appearance = facs_report_appearance(params),
  appearance_file = params$appearance_file,
  gate_fill = value_or(params$gate_fill, "#FFFFFF"),
  gate_alpha = value_or(params$gate_alpha, 0.08),
  label_color = value_or(params$label_color, "#FFFFFF"),
  label_size = value_or(params$label_size, 3.5),
  label_digits = value_or(params$label_digits, 1),
  gate_top_inset_fraction = value_or(params$gate_top_inset_fraction, 0.02)
)
facs_report_save_artifacts(
  analysis, bundle, params$output_analysis_rds,
  params$output_figure_bundle_rds, params$overwrite
)
```

## 4N pH3-positive populations

The outline is the exact FlowJo pH3-positive polygon clipped to the configured
G2/M DNA range. Its upper display edge is inset slightly from the y-axis
ceiling so the full outline remains visible. Each annotation is the percentage
of all Single Cell events that are both FlowJo pH3-positive and within the
configured G2/M range.

```{r}
bundle$pseudocolor
```

## Gate geometry and provenance

```{r}
unique(geometry[c(
  "prefix", "sample_id", "gate_name", "gate_path", "gate_type",
  "x_channel", "y_channel", "workspace"
)])
analysis$quantitation$ph3$sample_summary
analysis$provenance
```
