---
title: "Experimental model-derived EdU report (non-production)"
format: { html: { toc: true, embed-resources: true } }
execute: { warning: false, message: false }
params: { config: null, analysis_rds: null, model_gate_manifest: null }
---

```{r}
#| label: setup
#| include: false
library(facspseudocolor)
source(system.file("quarto", "_report-setup.R", package = "facspseudocolor"), local = TRUE)
if (is.null(params$model_gate_manifest)) stop("`model_gate_manifest` is required.")
analysis <- facs_report_input(params$config, params$analysis_rds)
model_gate_inputs <- validate_edu_model_gate_manifest(analysis, params$model_gate_manifest)
report <- build_edu_pseudocolor_output_contract(analysis)
```

> **EXPERIMENTAL MODEL-DERIVED NON-PRODUCTION.** Single Cells, G1, and EdU+
> were called from raw FCS events with immutable frozen models. No FlowJo
> workspace or FlowJo gate was used. This report
> does not establish biological ground truth, production performance, or broad
> generalizability.

The EdU classifier is the validated `PORTABLE_HGB_V1` export of
`EDU_POSITIVE_MODEL002`. It preserves the frozen model, feature order, and
calls without pickle deserialization or a scikit-learn/Python-version dependency.
The manifest's `derived_from_reference_sha256` records the validated predictor
source used to derive the embedded traversal; it is not an embedded-code digest.

## Every-sample DNA-versus-EdU panels

```{r}
for (prefix in names(report$panels)) print(report$panels[[prefix]])
```

## EdU-positive percentages

EdU positivity is a model call at the frozen 0.385 threshold, intersected by
direct event identity with each model-derived Single Cells population.

```{r}
print(report$positivity$panels$two_to_four_n)
print(report$positivity$panels$early_mid_late_s)
```

## EdU-positive intensity

```{r}
print(report$intensity$panels$all_computed_positive)
print(report$intensity$panels$early_mid_late_s)
```

## Model-gating provenance

```{r}
print(model_gate_inputs)
print(attr(model_gate_inputs, "model_artifacts"))
```

The G1 model was developed within expert-reference Single Cells. Applying it
after model-derived Single Cells changes its parent distribution; prior G1
validation metrics therefore do not validate this composed use. The older G1
rule also omits the second-marker axes used by the audited expert G1 gates.
For EdU experiments, model-called EdU+ events are excluded from the final G1
population before that G1 table is used for normalization or background fitting.
The operation also requires raw DNA-A to be at least 35% of the median positive
DNA-A among preliminary EdU-negative G1 candidates. This acquisition-specific
minimum adapts to detector scale and is recorded in the gating manifest.

The EdU model was validated only within expert-gated Single Cells. Its use here
within model-derived Single Cells changes the parent distribution and is an
explicitly approved experimental extension; its earlier validation metrics do
not validate this composed use. Frozen thresholds are Single Cells 0.205, G1
0.255, and EdU+ 0.385. These outputs must not be represented as biological
ground truth or production gating.

## Software used

```{r}
facs_report_software_used()
```
