---
title: "FACS analysis configurator"
subtitle: "Experimental, unfinished interface"
format:
  html:
    toc: false
    theme: cosmo
server: shiny
execute: { echo: false, warning: false, message: false }
---

> **Development status:** This configurator is experimental and unfinished. It
> is not part of the stable report workflow.

This guided first version creates configurations for **FlowJo-gated EdU
experiments**. It inspects the workspace, lets you include or exclude every
acquisition, records condition names and replicate membership, and downloads a
validated `config.yml`. No YAML editing is required. The required FlowJo-gated
CSV exports must already exist (or be created separately); this page does not
run or validate a FlowJo export operation.

Before running it, set RStudio's working directory to the experiment folder and
run `facspseudocolor::set_facs_configurator_directory()` in the Console.

```{r}
#| context: setup
library(shiny)
library(facspseudocolor)
library(xml2)

initial_data_dir <- Sys.getenv("FACS_CONFIG_DATA_DIR", unset = getwd())
if (!dir.exists(initial_data_dir)) initial_data_dir <- getwd()
initial_data_dir <- normalizePath(initial_data_dir, mustWork = TRUE)
workspace_paths <- list.files(
  initial_data_dir, pattern = "\\.wsp$", recursive = TRUE,
  full.names = TRUE, ignore.case = TRUE
)
workspace_choices <- stats::setNames(
  workspace_paths,
  substring(workspace_paths, nchar(initial_data_dir) + 2L)
)
config_paths <- list.files(
  initial_data_dir, pattern = "\\.ya?ml$", recursive = TRUE,
  full.names = TRUE, ignore.case = TRUE
)
config_choices <- stats::setNames(
  config_paths,
  substring(config_paths, nchar(initial_data_dir) + 2L)
)
```

```{r}
#| context: server
workspace_files <- reactiveVal(character())
sample_names <- reactiveVal(character())
loaded_config <- reactiveVal(NULL)
loaded_analysis <- reactiveVal(NULL)
spreadsheet_rows <- reactiveVal(NULL)
dna_height_confirmation <- reactiveVal(NULL)
sample_mapping_confirmation <- reactiveVal(NULL)
spreadsheet_mapping_confirmation <- reactiveVal(NULL)
spreadsheet_status <- reactiveVal("Load a FlowJo workspace, download the Excel template, edit it, and import it here.")
loaded_config_status <- reactiveVal("Select a configuration and click Load configuration.")
analysis_status <- reactiveVal("No analysis has been run in this session.")

inspect_workspace <- function(path) {
  document <- xml2::read_xml(path)
  datasets <- xml2::xml_find_all(document, "//*[local-name()='DataSet']")
  uris <- utils::URLdecode(xml2::xml_attr(datasets, "uri"))
  fcs <- basename(sub("^file:", "", uris))
  populations <- unique(xml2::xml_attr(
    xml2::xml_find_all(document, "//*[local-name()='Population']"), "name"
  ))
  channels <- unique(xml2::xml_attr(
    xml2::xml_find_all(document, "//*[local-name()='parameter']"), "name"
  ))
  list(fcs = fcs[nzchar(fcs)], populations = populations[nzchar(populations)],
       channels = channels[nzchar(channels)])
}

role_choices <- reactive({
  c("Experimental (background fitted from this acquisition)" = "experimental")
})

suffixes_for_mode <- function(mode) {
  switch(mode,
    edu = list(complete = "_single_cells.csv", g1 = "_g1.csv",
               edu_positive = "_edu_positive.csv"),
    poi = list(complete = "_single_cells.csv"),
    ph3 = list(complete = "_single_cells.csv", g1 = "_g1.csv",
               ph3_positive = "_ph3_positive.csv")
  )
}

safe_value <- function(value, fallback) {
  if (is.null(value) || !length(value) || is.na(value) || !nzchar(value)) fallback else value
}

guess_workspace_rows <- function(files) {
  stems <- tools::file_path_sans_ext(basename(files))
  conditions <- sub("(?i)[_ -]?rerun[0-9]*$", "", stems, perl = TRUE)
  conditions[!nzchar(conditions)] <- stems[!nzchar(conditions)]
  technical <- ave(seq_along(conditions), conditions, FUN = seq_along)
  data.frame(
    use = TRUE,
    order = seq_along(files),
    sample_name = conditions,
    fcs_file = basename(files),
    biological_replicate = "Biological replicate 1",
    technical_replicate = as.character(technical),
    role = "experimental",
    notes = "",
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

rows_from_spreadsheet <- function(tabular) {
  required <- c("use", "order", "sample_name", "fcs_file",
                "biological_replicate", "technical_replicate", "role")
  missing <- setdiff(required, names(tabular))
  if (length(missing)) stop("Spreadsheet is missing column(s): ", paste(missing, collapse = ", "))
  use <- tolower(trimws(as.character(tabular$use))) %in% c("true", "yes", "1", "y")
  tabular <- tabular[use, , drop = FALSE]
  if (!nrow(tabular)) stop("The spreadsheet has no rows marked for use.")
  if (any(!nzchar(trimws(tabular$sample_name))) || any(!nzchar(trimws(tabular$fcs_file)))) {
    stop("Every included row needs a sample_name and fcs_file.")
  }
  if (anyDuplicated(tolower(tabular$fcs_file))) stop("Each FCS file may appear only once.")
  valid_roles <- if (identical(input$plot_type, "poi")) c("experimental", "reference") else "experimental"
  if (any(!tolower(tabular$role) %in% valid_roles)) {
    stop("Role must be ", paste(valid_roles, collapse = " or "), " for this analysis type.")
  }
  lapply(seq_len(nrow(tabular)), function(i) list(
    include = TRUE,
    order = suppressWarnings(as.numeric(tabular$order[[i]])),
    prefix = tools::file_path_sans_ext(basename(tabular$fcs_file[[i]])),
    condition = trimws(as.character(tabular$sample_name[[i]])),
    replicate = trimws(as.character(tabular$biological_replicate[[i]])),
    technical_replicate = trimws(as.character(tabular$technical_replicate[[i]])),
    fcs = basename(tabular$fcs_file[[i]]),
    role = tolower(trimws(as.character(tabular$role[[i]])))
  ))
}

config_from_rows <- function(rows) {
  mode <- input$plot_type
  config <- list(
    plot_type = "edu",
    data_dir = "flowjo_gated_csv",
    dna_channel = "DNA content",
    target_channel = "EdU",
    target_name = "EdU",
    suffixes = suffixes_for_mode(mode),
    pseudocolor_signal = "background_subtracted",
    background_subtracted_offset = "auto",
    quant_signal = "background_subtracted",
    quantify_reference_normalized = isTRUE(input$reference_normalize),
    output_pdf = input$output_pdf,
    output_png = input$output_png,
    gating = list(mode = "flowjo"),
    replicates = build_replicates(rows, mode)
  )
  config$report <- list(
    all_events_dir = "all_events_csv",
    dna_height_channel = input$dna_height_channel,
    show_apex_comparison_toggle = isTRUE(input$show_apex_comparison_toggle),
    show_phase_gate_toggle = isTRUE(input$show_phase_gate_toggle),
    embed_pseudocolor_pdf_downloads = isTRUE(input$embed_pseudocolor_pdf_downloads)
  )
  if (isTRUE(input$reference_normalize) && !identical(mode, "ph3")) {
    config$quant_reference_condition <- input$normalization_reference
  }
  if (identical(mode, "ph3")) {
    boundaries <- c(input$ph3_g1_low, input$ph3_g1_high,
                    input$ph3_early_high, input$ph3_mid_high,
                    input$ph3_late_high, input$ph3_g2m_high)
    config$g1_x_range <- boundaries[c(1, 2)]
    config$s_phase_bins <- list(
      early = boundaries[c(2, 3)], mid = boundaries[c(3, 4)],
      late = boundaries[c(4, 5)]
    )
    config$g2m_x_range <- boundaries[c(5, 6)]
  }
  config
}

collect_sample_rows <- function() {
  names <- sample_names()
  if (!length(names)) stop("Create at least one sample name.")
  rows <- unlist(lapply(seq_along(names), function(i) {
    if (!isTRUE(input[[paste0("include_", i)]])) return(list())
    files <- input[[paste0("file_", i)]]
    if (is.null(files) || !length(files)) stop("Assign at least one FCS file to every included sample.")
    lapply(seq_along(files), function(j) list(
      include = TRUE,
      order = suppressWarnings(as.numeric(input[[paste0("order_", i)]])),
      prefix = tools::file_path_sans_ext(files[[j]]),
      condition = safe_value(input[[paste0("condition_", i)]], names[[i]]),
      replicate = safe_value(input[[paste0("bio_", i, "_", j)]], "Biological replicate 1"),
      technical_replicate = safe_value(input[[paste0("tech_", i, "_", j)]], as.character(j)),
      fcs = files[[j]],
      role = safe_value(input[[paste0("role_", i)]], "experimental")
    ))
  }), recursive = FALSE)
  if (!length(rows)) stop("Select at least one sample.")
  assigned <- vapply(rows, `[[`, character(1), "fcs")
  if (anyDuplicated(assigned)) stop("Each FCS file may be assigned to only one sample.")
  orders <- vapply(rows, `[[`, numeric(1), "order")
  orders[!is.finite(orders)] <- seq_along(rows)[!is.finite(orders)]
  rows[order(orders, seq_along(rows))]
}

build_replicates <- function(rows, mode) {
  replicate_names <- unique(vapply(rows, `[[`, character(1), "replicate"))
  lapply(replicate_names, function(replicate_name) {
    group <- rows[vapply(rows, function(x) identical(x$replicate, replicate_name), logical(1))]
    group <- group[order(
      vapply(group, `[[`, character(1), "technical_replicate"),
      vapply(group, `[[`, numeric(1), "order")
    )]
    out <- list(
      label = replicate_name,
      samples = lapply(group, function(x) list(
        label = x$condition, prefix = x$prefix, fcs = x$fcs,
        technical_replicate = x$technical_replicate
      ))
    )
    if (identical(mode, "poi")) {
      references <- vapply(group, function(x) identical(x$role, "reference"), logical(1))
      reference_conditions <- unique(vapply(group[references], `[[`, character(1), "condition"))
      if (length(reference_conditions) != 1L) {
        stop("Each biological replicate must name exactly one reference/background condition.")
      }
      out$reference <- reference_conditions[[1]]
      out <- out[c("label", "reference", "samples")]
    }
    out
  })
}

candidate_config <- reactive({
  config_from_rows(collect_sample_rows())
})

spreadsheet_config <- reactive({
  req(spreadsheet_rows())
  config_from_rows(rows_from_spreadsheet(spreadsheet_rows()))
})
```

::: {.panel-tabset}
## Create configuration

::: {.panel-tabset}
### 1. Experiment

```{r}
selectInput("plot_type", "Analysis type",
            choices = c("EdU versus DNA (FlowJo gates)" = "edu"))
fluidRow(
  column(12, textInput("data_dir_display", "RStudio working directory",
                       value = initial_data_dir, width = "100%"))
)
selectInput("workspace", "FlowJo workspace", choices = workspace_choices,
            width = "100%")
actionButton("load_workspace", "Load workspace", class = "btn-primary")
hr()
selectInput("dna_height_channel", "Raw DNA-H detector used in all-events CSVs",
            choices = character(), width = "50%")
checkboxInput("confirm_dna_height",
              "I confirmed the DNA-H detector against the experiment record",
              FALSE)
conditionalPanel(
  condition = "input.plot_type == 'ph3'",
  h4("Contiguous pH3 DNA phase boundaries"),
  fluidRow(
    column(2, numericInput("ph3_g1_low", "G1 lower", 900)),
    column(2, numericInput("ph3_g1_high", "G1 / Early", 1150)),
    column(2, numericInput("ph3_early_high", "Early / Mid", 1350)),
    column(2, numericInput("ph3_mid_high", "Mid / Late", 1600)),
    column(2, numericInput("ph3_late_high", "Late / G2M", 1850)),
    column(2, numericInput("ph3_g2m_high", "G2M upper", 2300))
  )
)
```

### 2. Samples

Loading the workspace creates one row per acquisition. Filename-based names and
replicate numbers are only starting suggestions; confirm every condition,
biological replicate, and technical replicate from the experiment record.

### Step 1: Review acquisition names

Keep one name per line. Give original acquisitions and reruns separate names so
their event data remain independent.

```{r}
textAreaInput("sample_names_text", "Sample names", rows = 8,
              placeholder = "NT original\nNT rerun 1\nIAA 2 h original\nIAA 2 h rerun")
actionButton("create_samples", "Create samples", class = "btn-primary")

h3("Step 2: Assign workspace files to each sample")
uiOutput("file_editor")

h3("Step 3: Identify biological and technical replicates")
uiOutput("acquisition_editor")

h3("Step 4: Define sample settings")
tags$p("Lower order numbers appear first. Clear Use to exclude a rerun or other acquisition. EdU background is fitted independently for every included acquisition.")
uiOutput("sample_editor")
checkboxInput("confirm_sample_mapping",
              "I confirmed every included acquisition, condition, and replicate assignment against the experiment record",
              FALSE)
```

### 3. Excel entry

This alternative starts from the FlowJo workspace. The workbook contains one
row per FCS acquisition and guesses sample names and technical-rerun numbers
from the filenames. Review all inferred values, especially biological
replicates and condition names, before importing it.

```{r}
downloadButton("download_excel_template", "Download workspace-filled Excel template")
br(); br()
fileInput("excel_config", "Import completed Excel template",
          accept = c(".xlsx"), width = "100%")
checkboxInput("confirm_spreadsheet_mapping",
              "I confirmed every included spreadsheet acquisition, condition, and replicate assignment against the experiment record",
              FALSE)
verbatimTextOutput("spreadsheet_status")
h4("Imported acquisitions")
tableOutput("spreadsheet_preview")
downloadButton("download_spreadsheet_config", "Download validated YAML from Excel")
```

### 4. Output

```{r}
checkboxInput("reference_normalize",
              "Normalize background-subtracted medians to a reference condition",
              value = FALSE)
conditionalPanel(
  condition = "input.reference_normalize && input.plot_type != 'ph3'",
  selectInput("normalization_reference", "Post-subtraction normalization reference",
              choices = character())
)
fluidRow(
  column(6, textInput("output_pdf", "Output PDF", "results/pseudocolor_panels.pdf")),
  column(6, textInput("output_png", "Output PNG", "results/pseudocolor_panels.png"))
)
checkboxInput("show_apex_comparison_toggle", "Include apex comparison control", TRUE)
checkboxInput("show_phase_gate_toggle", "Include cell-cycle gate-box control", TRUE)
checkboxInput("embed_pseudocolor_pdf_downloads", "Include vector-PDF downloads", TRUE)
actionButton("validate_config", "Validate configuration", class = "btn-primary")
downloadButton("download_config", "Download config.yml")
br(); br()
verbatimTextOutput("validation_status")
h4("YAML preview")
verbatimTextOutput("yaml_preview")
h4("Report command")
verbatimTextOutput("render_command")
```
:::

## Load and run

Choose a saved configuration from the current RStudio working directory. Loading
is read-only: the selected YAML file is not changed.

```{r}
fluidRow(
  column(9, selectInput("existing_config", "Saved configuration",
                        choices = config_choices, width = "100%")),
  column(3, br(), actionButton("refresh_configs", "Refresh list"))
)
actionButton("load_existing_config", "Load configuration", class = "btn-primary")
actionButton("run_existing_config", "Run analysis", class = "btn-success")
br(); br()
verbatimTextOutput("loaded_config_status")
h4("Configuration summary")
tableOutput("loaded_config_summary")
h4("Acquisitions")
tableOutput("loaded_manifest")
```

## Results

Results appear after **Run analysis** completes.

::: {.panel-tabset}
### Pseudocolor plots

```{r}
plotOutput("result_pseudocolor", height = "900px")
```

### EdU phase barplots

```{r}
plotOutput("result_phase_bars", height = "650px")
```

### Background diagnostics

```{r}
plotOutput("result_background_diagnostics", height = "900px")
```

### Quantitation tables

```{r}
h4("Phase medians")
tableOutput("result_phase_table")
h4("Whole-population medians")
tableOutput("result_whole_table")
```
:::
:::

```{r}
#| context: server
observeEvent(input$plot_type, {
  updateTextInput(session, "target_name", value = switch(
    input$plot_type, edu = "EdU", poi = "Target", ph3 = "pH3"
  ))
})

observeEvent(input$load_workspace, {
  if (is.null(input$workspace) || !file.exists(input$workspace)) {
    showNotification("Select an existing FlowJo .wsp file.", type = "error")
    return()
  }
  info <- tryCatch(inspect_workspace(input$workspace), error = identity)
  if (inherits(info, "error")) {
    showNotification(conditionMessage(info), type = "error")
    return()
  }
  workspace_files(info$fcs)
  suggested_names <- make.unique(tools::file_path_sans_ext(basename(info$fcs)))
  sample_names(suggested_names)
  updateTextAreaInput(session, "sample_names_text",
                      value = paste(suggested_names, collapse = "\n"))
  updateSelectInput(session, "normalization_reference",
                    choices = suggested_names,
                    selected = if (length(suggested_names)) suggested_names[[1]] else character())
  dna_height_default <- if ("FL2-H" %in% info$channels) "FL2-H" else ""
  updateSelectInput(session, "dna_height_channel",
                    choices = c("Choose a detector" = "", info$channels),
                    selected = dna_height_default)
  dna_height_confirmation(NULL)
  sample_mapping_confirmation(NULL)
  updateCheckboxInput(session, "confirm_dna_height", value = FALSE)
  updateCheckboxInput(session, "confirm_sample_mapping", value = FALSE)
})

observeEvent(input$create_samples, {
  names <- trimws(unlist(strsplit(input$sample_names_text, "\\r?\\n")))
  names <- names[nzchar(names)]
  if (!length(names)) {
    showNotification("Enter at least one sample name.", type = "error")
    return()
  }
  if (anyDuplicated(names)) {
    showNotification("Sample names must be unique.", type = "error")
    return()
  }
  sample_names(names)
  updateSelectInput(session, "normalization_reference", choices = names,
                    selected = names[[1]])
})

output$file_editor <- renderUI({
  names <- sample_names()
  files <- workspace_files()
  if (!length(names)) return(tags$em("Create sample names first."))
  if (!length(files)) return(tags$em("Load a FlowJo workspace first."))
  tagList(lapply(seq_along(names), function(i) {
    selected_file <- if (i <= length(files)) files[[i]] else character()
    selectizeInput(paste0("file_", i), names[[i]], choices = files,
                   selected = selected_file, multiple = TRUE,
                   options = list(plugins = list("remove_button")))
  }))
})

output$acquisition_editor <- renderUI({
  names <- sample_names()
  if (!length(names)) return(tags$em("Create sample names first."))
  blocks <- lapply(seq_along(names), function(i) {
    files <- input[[paste0("file_", i)]]
    if (is.null(files) || !length(files)) return(NULL)
    tagList(
      tags$h5(names[[i]]),
      lapply(seq_along(files), function(j) fluidRow(
        column(4, tags$p(basename(files[[j]]))),
        column(4, textInput(paste0("bio_", i, "_", j),
                            "Biological replicate", "Biological replicate 1")),
        column(4, textInput(paste0("tech_", i, "_", j),
                            "Technical replicate", as.character(j)))
      ))
    )
  })
  if (!length(Filter(Negate(is.null), blocks))) tags$em("Assign files first.") else tagList(blocks)
})

output$sample_editor <- renderUI({
  names <- sample_names()
  if (!length(names)) return(tags$em("Create sample names first."))
  tagList(
    tags$div(class = "row", style = "font-weight:600; margin-bottom:6px;",
      tags$div(class = "col-sm-1", "Use"), tags$div(class = "col-sm-1", "Order"),
      tags$div(class = "col-sm-5", "Display name"),
      tags$div(class = "col-sm-5", "Sample role")
    ),
    lapply(seq_along(names), function(i) {
      fluidRow(
        column(1, checkboxInput(paste0("include_", i), NULL, TRUE)),
        column(1, numericInput(paste0("order_", i), NULL, i, min = 1, step = 1)),
        column(5, textInput(paste0("condition_", i), NULL, names[[i]])),
        column(5, selectInput(paste0("role_", i), NULL, role_choices()))
      )
    })
  )
})

confirm_config_inputs <- function() {
  if (!is.character(input$dna_height_channel) ||
      length(input$dna_height_channel) != 1L ||
      !nzchar(input$dna_height_channel)) {
    stop("Choose the DNA-H detector; no detector fallback is allowed.")
  }
  if (!isTRUE(input$confirm_dna_height)) {
    stop("Confirm the DNA-H detector against the experiment record.")
  }
  if (!identical(dna_height_confirmation(), input$dna_height_channel)) {
    stop("The DNA-H detector changed after confirmation. Clear and re-check its confirmation.")
  }
  if (!isTRUE(input$confirm_sample_mapping)) {
    stop("Confirm every included acquisition, condition, and replicate assignment against the experiment record.")
  }
  current_mapping <- yaml::as.yaml(collect_sample_rows())
  if (!identical(sample_mapping_confirmation(), current_mapping)) {
    stop("The acquisition mapping changed after confirmation. Clear and re-check its confirmation.")
  }
  invisible(TRUE)
}

observeEvent(input$confirm_dna_height, {
  dna_height_confirmation(if (isTRUE(input$confirm_dna_height))
    input$dna_height_channel else NULL)
})

observeEvent(input$confirm_sample_mapping, {
  if (!isTRUE(input$confirm_sample_mapping)) {
    sample_mapping_confirmation(NULL)
    return()
  }
  signature <- tryCatch(yaml::as.yaml(collect_sample_rows()), error = identity)
  if (inherits(signature, "error")) {
    sample_mapping_confirmation(NULL)
    updateCheckboxInput(session, "confirm_sample_mapping", value = FALSE)
    showNotification(conditionMessage(signature), type = "error")
  } else {
    sample_mapping_confirmation(signature)
  }
})

confirm_spreadsheet_inputs <- function() {
  if (!isTRUE(input$confirm_dna_height) ||
      !identical(dna_height_confirmation(), input$dna_height_channel)) {
    stop("Confirm the current DNA-H detector against the experiment record.")
  }
  if (!isTRUE(input$confirm_spreadsheet_mapping)) {
    stop("Confirm every included spreadsheet acquisition, condition, and replicate assignment against the experiment record.")
  }
  current_mapping <- yaml::as.yaml(rows_from_spreadsheet(spreadsheet_rows()))
  if (!identical(spreadsheet_mapping_confirmation(), current_mapping)) {
    stop("The imported spreadsheet mapping changed after confirmation. Re-import it and confirm again.")
  }
  invisible(TRUE)
}

observeEvent(input$confirm_spreadsheet_mapping, {
  if (!isTRUE(input$confirm_spreadsheet_mapping)) {
    spreadsheet_mapping_confirmation(NULL)
    return()
  }
  signature <- tryCatch(
    yaml::as.yaml(rows_from_spreadsheet(spreadsheet_rows())),
    error = identity
  )
  if (inherits(signature, "error")) {
    spreadsheet_mapping_confirmation(NULL)
    updateCheckboxInput(session, "confirm_spreadsheet_mapping", value = FALSE)
    showNotification(conditionMessage(signature), type = "error")
  } else {
    spreadsheet_mapping_confirmation(signature)
    checked <- tryCatch({
      confirm_spreadsheet_inputs()
      validate_facs_config(spreadsheet_config())
    }, error = identity)
    spreadsheet_status(if (inherits(checked, "error"))
      paste("Configuration needs attention:", conditionMessage(checked)) else
      "Spreadsheet configuration and confirmations are valid.")
  }
})

validation <- eventReactive(input$validate_config, {
  tryCatch({
    confirm_config_inputs()
    validated <- validate_facs_config(candidate_config())
    list(message = paste("Configuration is valid:",
      nrow(build_sample_manifest(validated)), "samples across",
      length(validated$replicates), "replicate(s)."))
  }, error = function(e) list(message = conditionMessage(e)))
})

output$validation_status <- renderText(validation()$message)
output$yaml_preview <- renderText({
  input$validate_config
  isolate(tryCatch(yaml::as.yaml(candidate_config()),
                   error = function(e) conditionMessage(e)))
})
output$download_config <- downloadHandler(
  filename = function() "config.yml",
  content = function(file) {
    confirm_config_inputs()
    config <- candidate_config()
    validate_facs_config(config)
    yaml::write_yaml(config, file)
  }
)

output$render_command <- renderText({
  report_path <- system.file(
    "quarto", "facs_edu_standard_v2.qmd",
    package = "facspseudocolor"
  )
  paste("quarto render", shQuote(report_path), "-P", shQuote("config=config.yml"))
})

output$download_excel_template <- downloadHandler(
  filename = function() paste0("facs_samples_", input$plot_type, ".xlsx"),
  content = function(file) {
    if (!requireNamespace("openxlsx", quietly = TRUE)) {
      stop("Install the R package 'openxlsx' to create Excel templates.")
    }
    files <- workspace_files()
    if (!length(files)) stop("Load a FlowJo workspace before downloading the template.")
    rows <- guess_workspace_rows(files)
    workbook <- openxlsx::createWorkbook()
    openxlsx::addWorksheet(workbook, "Instructions", gridLines = FALSE)
    openxlsx::addWorksheet(workbook, "Samples", gridLines = FALSE)
    instructions <- data.frame(
      Field = c("Purpose", "Autopopulated", "Review carefully", "EdU role",
                "Technical replicates", "Biological replicates"),
      Guidance = c(
        "One Samples row represents one FCS acquisition.",
        "FCS filenames come from the selected FlowJo workspace. Sample names and technical numbers are filename-based guesses.",
        "Confirm sample_name, biological_replicate, technical_replicate, role, and order.",
        "Use experimental for every acquisition; EdU background is fitted within each acquisition.",
        "Acquisitions sharing sample_name and biological_replicate are processed separately and then averaged.",
        "Use different biological_replicate labels for independently grown or treated experiments."
      ), check.names = FALSE
    )
    title_style <- openxlsx::createStyle(fontSize = 16, textDecoration = "bold",
                                          fontColour = "#FFFFFF", fgFill = "#176B87")
    header_style <- openxlsx::createStyle(textDecoration = "bold",
                                           fontColour = "#FFFFFF", fgFill = "#2E7D8F")
    input_style <- openxlsx::createStyle(fgFill = "#EAF4F6")
    openxlsx::writeData(workbook, "Instructions", "FACS configuration workbook", startRow = 1)
    openxlsx::addStyle(workbook, "Instructions", title_style, rows = 1, cols = 1:2, gridExpand = TRUE)
    openxlsx::writeDataTable(workbook, "Instructions", instructions, startRow = 3,
                             tableStyle = "TableStyleMedium2")
    openxlsx::setColWidths(workbook, "Instructions", cols = 1, widths = 24)
    openxlsx::setColWidths(workbook, "Instructions", cols = 2, widths = 80)
    openxlsx::setRowHeights(workbook, "Instructions", rows = 4:(nrow(instructions) + 3), heights = 32)
    openxlsx::addStyle(workbook, "Instructions",
                       openxlsx::createStyle(wrapText = TRUE, valign = "top"),
                       rows = 3:(nrow(instructions) + 3), cols = 1:2, gridExpand = TRUE)
    openxlsx::writeDataTable(workbook, "Samples", rows, tableStyle = "TableStyleMedium2")
    openxlsx::addStyle(workbook, "Samples", header_style, rows = 1,
                       cols = seq_len(ncol(rows)), gridExpand = TRUE, stack = TRUE)
    openxlsx::addStyle(workbook, "Samples", input_style, rows = 2:(nrow(rows) + 1),
                       cols = c(1:3, 5:8), gridExpand = TRUE)
    openxlsx::freezePane(workbook, "Samples", firstRow = TRUE)
    openxlsx::setColWidths(workbook, "Samples", cols = 1:8,
                           widths = c(9, 9, 28, 34, 25, 20, 16, 30))
    openxlsx::dataValidation(workbook, "Samples", cols = 1, rows = 2:(nrow(rows) + 1),
                             type = "list", value = "'TRUE,FALSE'")
    roles <- if (identical(input$plot_type, "poi")) "'experimental,reference'" else "'experimental'"
    openxlsx::dataValidation(workbook, "Samples", cols = 7, rows = 2:(nrow(rows) + 1),
                             type = "list", value = roles)
    openxlsx::saveWorkbook(workbook, file, overwrite = TRUE)
  }
)

observeEvent(input$excel_config, {
  spreadsheet_mapping_confirmation(NULL)
  updateCheckboxInput(session, "confirm_spreadsheet_mapping", value = FALSE)
  if (!requireNamespace("readxl", quietly = TRUE)) {
    spreadsheet_rows(NULL)
    spreadsheet_status("Install the R package 'readxl' to import Excel templates.")
    return()
  }
  result <- tryCatch(
    as.data.frame(readxl::read_excel(input$excel_config$datapath, sheet = "Samples"),
                  stringsAsFactors = FALSE),
    error = identity
  )
  if (inherits(result, "error")) {
    spreadsheet_rows(NULL)
    spreadsheet_status(paste("Excel import error:", conditionMessage(result)))
    return()
  }
  spreadsheet_rows(result)
  checked <- tryCatch(validate_facs_config(spreadsheet_config()), error = identity)
  spreadsheet_status(if (inherits(checked, "error")) {
    paste("Imported, but configuration needs attention:", conditionMessage(checked))
  } else {
    paste("Imported", nrow(result),
          "acquisition row(s). Confirm the spreadsheet mapping before download.")
  })
})

output$spreadsheet_status <- renderText(spreadsheet_status())
output$spreadsheet_preview <- renderTable({
  rows <- spreadsheet_rows()
  req(rows)
  rows
}, striped = TRUE, bordered = TRUE, spacing = "s")

output$download_spreadsheet_config <- downloadHandler(
  filename = function() paste0("facs_config_", input$plot_type, "_from_excel.yml"),
  content = function(file) {
    confirm_spreadsheet_inputs()
    config <- validate_facs_config(spreadsheet_config())
    yaml::write_yaml(config, file)
  }
)

discover_configs <- function() {
  paths <- list.files(
    initial_data_dir, pattern = "\\.ya?ml$", recursive = TRUE,
    full.names = TRUE, ignore.case = TRUE
  )
  stats::setNames(paths, substring(paths, nchar(initial_data_dir) + 2L))
}

observeEvent(input$refresh_configs, {
  choices <- discover_configs()
  updateSelectInput(session, "existing_config", choices = choices,
                    selected = if (length(choices)) unname(choices[[1]]) else character())
  showNotification(paste(length(choices), "configuration file(s) found."),
                   type = "message")
})

observeEvent(input$load_existing_config, {
  path <- input$existing_config
  if (is.null(path) || !length(path) || !file.exists(path)) {
    loaded_config(NULL)
    loaded_config_status("Select an existing YAML configuration.")
    return()
  }
  result <- tryCatch(read_facs_config(path), error = identity)
  if (inherits(result, "error")) {
    loaded_config(NULL)
    loaded_config_status(paste("Configuration error:", conditionMessage(result)))
    return()
  }
  loaded_config(result)
  loaded_analysis(NULL)
  manifest <- build_sample_manifest(result)
  loaded_config_status(paste(
    "Configuration is valid:", basename(path), "contains", nrow(manifest),
    "acquisitions across", length(unique(manifest$replicate)),
    "biological replicate(s)."
  ))
})

observeEvent(input$run_existing_config, {
  config <- loaded_config()
  if (is.null(config)) {
    analysis_status("Load a valid configuration before running the analysis.")
    loaded_config_status(analysis_status())
    return()
  }
  loaded_config_status("Running analysis. This can take a few moments...")
  result <- tryCatch(
    withProgress(message = "Running FACS analysis", value = 0.2, {
      analysis <- analyze_facs_experiment(config)
      incProgress(0.8, detail = "Preparing plots and tables")
      analysis
    }),
    error = identity
  )
  if (inherits(result, "error")) {
    loaded_analysis(NULL)
    analysis_status(paste("Analysis error:", conditionMessage(result)))
  } else {
    loaded_analysis(result)
    analysis_status(paste(
      "Analysis complete:", nrow(result$sample_manifest),
      "acquisitions processed with", length(result$warnings), "warning(s)."
    ))
  }
  loaded_config_status(analysis_status())
})

output$loaded_config_status <- renderText(loaded_config_status())

output$loaded_config_summary <- renderTable({
  config <- loaded_config()
  req(config)
  manifest <- build_sample_manifest(config)
  data.frame(
    Setting = c("Analysis", "Target", "DNA channel", "Signal channel",
                "Acquisitions", "Conditions", "Population data folder"),
    Value = c(config$plot_type, config$target_name, config$dna_channel,
              config$target_channel, nrow(manifest),
              length(unique(manifest$condition)), config$data_dir),
    check.names = FALSE
  )
}, striped = TRUE, bordered = TRUE, spacing = "s")

output$loaded_manifest <- renderTable({
  config <- loaded_config()
  req(config)
  manifest <- build_sample_manifest(config)
  keep <- intersect(c("replicate", "technical_replicate", "condition", "prefix"),
                    names(manifest))
  manifest[, keep, drop = FALSE]
}, striped = TRUE, bordered = TRUE, spacing = "s")

output$result_pseudocolor <- renderPlot({
  analysis <- loaded_analysis()
  req(analysis)
  plot_pseudocolor_panels(analysis)
}, res = 110)

output$result_phase_bars <- renderPlot({
  analysis <- loaded_analysis()
  req(analysis)
  validate(need(analysis$config$plot_type != "ph3",
                "EdU/POI phase-signal bars are not available in pH3 mode."))
  plot_facs_phase_signal(analysis, geometry = "bar",
                         signal = "background_subtracted")
}, res = 110)

output$result_background_diagnostics <- renderPlot({
  analysis <- loaded_analysis()
  req(analysis)
  plots <- lapply(analysis$models, function(model) model$diagnostic_plot)
  plots <- Filter(function(x) inherits(x, "ggplot"), plots)
  validate(need(length(plots), "No background diagnostic plots are available."))
  assemble_facs_panels(plots, ncol = min(3L, length(plots)))
}, res = 110)

output$result_phase_table <- renderTable({
  analysis <- loaded_analysis()
  req(analysis)
  analysis$quantitation$phase_medians
}, striped = TRUE, bordered = TRUE, spacing = "s")

output$result_whole_table <- renderTable({
  analysis <- loaded_analysis()
  req(analysis)
  analysis$quantitation$whole_medians
}, striped = TRUE, bordered = TRUE, spacing = "s")
```
